---
output:
  pdf_document: default
  word_document: default
  html_document: default
---
# AUTHOR: 42235_WURAOLA_OKUWOBI #

## QUESTION 1 ##

*1a. IMPORT DATA INTO R*
```{r 1a. IMPORT DATA INTO R}
library(readxl)
monthly_stock_returns <- read_excel("data_assignment1.xlsx", sheet = "stock_monthly")
```

#1b. TRANSFORM THE SIMPLE RETURNS INTO LOG RETURNS#
```{r #1b. TRANSFORM THE SIMPLE RETURNS INTO LOG RETURNS#}
monthly_stock_returns$COCACOLA1 <- c(0, diff(log(1+monthly_stock_returns$COCACOLA), lag = 1))
monthly_stock_returns$GE1 <- c(0, diff(log(1+monthly_stock_returns$GE), lag = 1))
monthly_stock_returns$IBM1 <- c(0, diff(log(1+monthly_stock_returns$IBM), lag = 1))
monthly_stock_returns$VWRET1 <- c(0, diff(log(1+monthly_stock_returns$VWRET), lag = 1))
monthly_stock_returns$EWRET1 <- c(0, diff(log(1+monthly_stock_returns$EWRET), lag = 1))
```

#1c. PLOT THE SIMPLE AND LOG RETURNS FOR ALL THE SERIES#
```{r #1c. PLOT THE SIMPLE AND LOG RETURNS FOR ALL THE SERIES#}
plot(monthly_stock_returns$COCACOLA, monthly_stock_returns$COCACOLA1)
plot(monthly_stock_returns$GE, monthly_stock_returns$GE1)
plot(monthly_stock_returns$IBM, monthly_stock_returns$IBM1)
plot(monthly_stock_returns$VWRET, monthly_stock_returns$VWRET1)
plot(monthly_stock_returns$EWRET, monthly_stock_returns$EWRET1)
```

#1d. SCATTERPLOT OF ALL THE LOG RETURNS SERIES#
```{r #1d. SCATTERPLOT OF ALL THE LOG RETURNS SERIES#}
monthly_stock_returns_log <- monthly_stock_returns[,c(7:11)]
plot(monthly_stock_returns_log)
```

#1e. SAMPLE MEAN, VARIANCE, SKEWNESS, EXCESS KURTOSIS, MINIMUM AND MAXIMUM OF LOG RETURNS#
```{r #1e. SAMPLE MEAN, VARIANCE, SKEWNESS, EXCESS KURTOSIS, MINIMUM AND MAXIMUM OF LOG RETURNS#}
library(moments)
COCACOLA_log <- c(mean(monthly_stock_returns$COCACOLA1), var(monthly_stock_returns$COCACOLA1), skewness(monthly_stock_returns$COCACOLA1), kurtosis(monthly_stock_returns$COCACOLA1), min(monthly_stock_returns$COCACOLA1), max(monthly_stock_returns$COCACOLA1))
GE_log <- c(mean(monthly_stock_returns$GE1), var(monthly_stock_returns$GE1), skewness(monthly_stock_returns$GE1), kurtosis(monthly_stock_returns$GE1), min(monthly_stock_returns$GE1), max(monthly_stock_returns$GE1))
IBM_log <- c(mean(monthly_stock_returns$IBM1), var(monthly_stock_returns$IBM1), skewness(monthly_stock_returns$IBM1), kurtosis(monthly_stock_returns$IBM1), min(monthly_stock_returns$IBM1), max(monthly_stock_returns$IBM1))
VWRET_log <- c(mean(monthly_stock_returns$VWRET1), var(monthly_stock_returns$VWRET1), skewness(monthly_stock_returns$VWRET1), kurtosis(monthly_stock_returns$VWRET1), min(monthly_stock_returns$VWRET1), max(monthly_stock_returns$VWRET1))
EWRET_log <- c(mean(monthly_stock_returns$EWRET1), var(monthly_stock_returns$EWRET1), skewness(monthly_stock_returns$EWRET1), kurtosis(monthly_stock_returns$EWRET1), min(monthly_stock_returns$EWRET1), max(monthly_stock_returns$EWRET1))

summary_df <- data.frame(COCACOLA_log, GE_log, IBM_log, VWRET_log, EWRET_log)
row.names(summary_df) <- c("MEAN", "VARIANCE", "SKEWNESS", "EXCESS KURTOSIS", "MINIMUM", "MAXIMUM")
summary_df
```

#1f. AT 5% SIGNIFICANCE LEVEL, ARE SAMPLE MEAN, SKEWNESS AND EXCESS KURTOSIS OF LOG RETURNS STATISTICALLY DIFFERENT FROM ZERO? #
```{r #1f. AT 5% SIGNIFICANCE LEVEL, ARE SAMPLE MEAN, SKEWNESS AND EXCESS KURTOSIS OF LOG RETURNS STATISTICALLY DIFFERENT FROM ZERO? #}
# SAMPLE MEAN TEST

mean_COCACOLA1 <- t.test(monthly_stock_returns_log$COCACOLA1)
mean_GE1 <- t.test(monthly_stock_returns_log$GE1)
mean_IBM1 <- t.test(monthly_stock_returns_log$IBM1)
mean_VWRET1 <- t.test(monthly_stock_returns_log$VWRET1)
mean_EWRET1 <- t.test(monthly_stock_returns_log$EWRET1)

# Answer: Since the p-value > 5%, at 5% significance level we do not reject the null hypothesis that the sample means of the log returns are statistically equL zero. This means that the true means of the log returns of the variables (COCACOLA, IBM, GE, VWRET AND EWRET) are equal to zero.

# SKEWNESS TEST #
library(moments)
skew_COCACOLA1 <- skewness(monthly_stock_returns_log$COCACOLA1)/sqrt(6/588)
skew_GE1 <- skewness(monthly_stock_returns_log$GE1)/sqrt(6/588)
skew_IBM1 <- skewness(monthly_stock_returns_log$IBM1)/sqrt(6/588)
skew_VWRET1 <- skewness(monthly_stock_returns_log$VWRET1)/sqrt(6/588)
skew_EWRET1 <- skewness(monthly_stock_returns_log$EWRET1)/sqrt(6/588)

#Answer: Since the absolute value from the result of the skewness test > 1.96, at 5% significance level we reject the null hypothesis that the skewness of the log returns are equal to zero. This means that the log returns of the variables (COCACOLA, IBM, GE, VWRET AND EWRET) do not follow normal distribution.#

# KURTOSIS TEST #

kurtosis_COCACOLA1 <- (kurtosis(monthly_stock_returns_log$COCACOLA1)-3)/sqrt(24/588)
kurtosis_GE1 <- (kurtosis(monthly_stock_returns_log$GE1)-3)/sqrt(24/588)
kurtosis_IBM1 <- (kurtosis(monthly_stock_returns_log$IBM1)-3)/sqrt(24/588)
kurtosis_VWRET1 <- (kurtosis(monthly_stock_returns_log$VWRET1)-3)/sqrt(24/588)
kurtosis_EWRET1 <- (kurtosis(monthly_stock_returns_log$EWRET1)-3)/sqrt(24/588)

#Answer: Since the value from the result of the kurtosis test > 1.96, at 5% significance level we reject the null hypothesis that the kurtosis of the log returns are equal to zero. This means that the log returns of the variables (COCACOLA, IBM, GE, VWRET AND EWRET) do not follow normal distribution.
```

#1g. HISTOGRAM OF RETURNS IN COMPARISON TO NORMAL AND STUDENT DISTRIBUTIONS #
```{r #1g. HISTOGRAM OF RETURNS IN COMPARISON TO NORMAL AND STUDENT DISTRIBUTIONS #}

#JARQUE - BERA TEST AT THE 5% SIGNIFICANCE LEVEL #

library(tseries)

# Answer: Since the p-values from the result of the jarque-bera tests for the returns and log returns of COCACOLA, GE, IBM, VWRET and EWRET are less than 5%, then we reject the null hypothesis and conclude that the returns and log returns do not follow a normal distribution.
```

#1H. CAPM REGRESSION FOR COCACOLA, GE AND IBM. #
```{r #1H. CAPM REGRESSION FOR COCACOLA, GE AND IBM. #}
COCACOLA_FIT <- lm(COCACOLA ~ VWRET, data = monthly_stock_returns)
GE_FIT <- lm(GE ~ VWRET, data = monthly_stock_returns)
IBM_FIT <- lm(IBM ~ VWRET, data = monthly_stock_returns)
summary(COCACOLA_FIT)
summary(GE_FIT)
summary(IBM_FIT)

# Answer: Please note that the VWRET was selected as the proxy for market returns because the market portfolio is a value-weighted portfolio of all securities traded in the market. The alphas α for COCACOLA, GE and IBM are 0.006496, 0.0004826 and 0.002762 respectively while the betas β for COCACOLA, GE and IBM are 0.716446, 1.1199868 and 0.879652 respectively. The p-value of the β are less than 1.96, hence at 5% significance levels w edo not reject the null hypothesis and conclude that are statistically significant. 

```

## QUESTION 2 ##

#2a. IMPORT DATA INTO R #
```{r #2a. IMPORT DATA INTO R #}
library("haven")

pension_df <-read_dta("pension.dta")
```

#2b. HOW MANY SINGLE PERSON HOUSEHOLDS ARE THERE IN THE DATASET #
```{r #2b. HOW MANY SINGLE PERSON HOUSEHOLDS ARE THERE IN THE DATASET #}
library("dplyr")

no_of_single_person_households <- length(which(pension_df$marr == 0))

no_of_single_person_households
```

#2c. OLS ESTIMATION OF THE MODEL TO REFLECT RESULTS OF SINGLE-PERSON HOUSEHOLDS
```{r}
OLS_estimatea <- lm(nettfa ~ inc + age, data = pension_df)

summary(OLS_estimatea)

OLS_estimate_single <- lm(nettfa ~ inc + age, data = subset(pension_df, marr == 0))

summary(OLS_estimate_single)
```

#2d. MEANING OF INTERCEPT FROM THE REGRESSION#
```{r}
# Answer: The intercept shows that single people have negative net financial wealth of -43-97943 if income and age equals zero. However, this negative net financial wealth position is worse for married people at an intercept of -60-69654.
```

#2g. OLS ESTIMATION OF THE MODEL#
```{r}
OLS_estimate1 <- lm(nettfa ~ inc, data = pension_df)

summary(OLS_estimate1)
```

#2h. OLS ESTIMATION OF THE MODEL
```{r}
OLS_estimate2 <- lm(nettfa ~ inc + age + incsq + agesq + fsize, data = pension_df)

summary(OLS_estimate2)
```

#2i. F-TEST OF THE RESTRICTION
```{r}
library(car)
nullhyp <- c("incsq", "agesq")
linearHypothesis(OLS_estimate2, nullhyp)

#Answer: The probability value of the f-test is less than 1%, hence we reject the null hypothesis at 1% significance level and conclude that the coefficients of incsq and agesq do not equal zero.
```

#2j. RE-ESTIMATION OF THE MODEL (OLS_estimate2) AFTER RESCALING inc BY DIVIDING IT BY 10#
```{r #2j. RE-ESTIMATION OF THE MODEL (OLS_estimate2) AFTER RESCALING inc BY DIVIDING IT BY 10#}
pension_df$inc_adjusted <- pension_df$inc/10

OLS_estimate2_adj <- lm(nettfa ~ inc_adjusted + age + I(inc^2) + I(age^2) + fsize, data = pension_df)

summary(OLS_estimate2_adj)
```

#2k. TESTING THE MODEL (OLS_estimate) FOR HETEROSKEDASTICITY USING THE BREUSCH-PAGAN TEST #
```{r #2k. TESTING THE MODEL (OLS_estimate) FOR HETEROSKEDASTICITY USING THE BREUSCH-PAGAN TEST #}
library(lmtest)
bptest(OLS_estimatea)

#Answer: the p-value is less than the significance level at 1% or 5%, hence we reject the null hypothesis and conclude that heteroskedasticity is present #

```

#2l. ESTIMATING THE MODEL (OLS_estimate) WITH HETEROSKEDASTICITY-ROBUST (e.g. WHITE) STANDARD ERRORS #
```{r #2l. ESTIMATING THE MODEL (OLS_estimate) WITH HETEROSKEDASTICITY-ROBUST (e.g. WHITE) STANDARD ERRORS #}
library(lmtest)
library(sandwich)
coeftest(OLS_estimatea, vcov = sandwich)
```

#2m. RE-ESTIMATING THE MODEL (OLS_estimate) AFTER STANDARDISING ALL VARIABLES IN THE MODEL. INTERPRETE COEFFICIENTS ON inc AND age #
```{r #2m. RE-ESTIMATING THE MODEL (OLS_estimate) AFTER STANDARDISING ALL VARIABLES IN THE MODEL. INTERPRETE COEFFICIENTS ON inc AND age #}
OLS_estimatea <- lm(nettfa ~ inc + age, data = pension_df)
pension_df$resi <- OLS_estimatea$residuals
varfunc.ols <- lm(log(resi^2) ~ log(inc) + log(age), data = pension_df)
pension_df$varfunc <- exp(varfunc.ols$fitted.values)
OLS_estimatea_gls <- lm(nettfa ~ inc + age, weights = 1/sqrt(varfunc), data = pension_df)

summary(OLS_estimatea)
summary(OLS_estimatea_gls)
```

## QUESTION 3 ##

# IMPORT DATA INTO R #
```{r}
library("haven")
library("sandwich")
library(foreign)
library("lmtest")
library("zoo")
library(dplyr)

ceo_salary_df <-read_dta("ceo_salary.dta")
```

#3a. Estimation of the model using OLS
```{r}
OLS_estimate_salary <- lm(lsalary ~ lsales + lmktval + ceoten + ceotensq, data = ceo_salary_df)

summary(OLS_estimate_salary)
```

#3b. COUNT, MEAN, STANDARD DEVIATION, MINIMUM AND MAXIMUM OF EXPLANATORY VARIABLES. EXPLAIN REASON FOR NATURAL LOG
```{r #3b. COUNT, MEAN, STANDARD DEVIATION, MINIMUM AND MAXIMUM OF EXPLANATORY VARIABLES. EXPLAIN REASON FOR NATURAL LOG}
lsales. <- c(length(ceo_salary_df$lsales), mean(ceo_salary_df$lsales), sd(ceo_salary_df$lsales), min(ceo_salary_df$lsales), max(ceo_salary_df$lsales))
lmktval. <- c(length(ceo_salary_df$lmktval), mean(ceo_salary_df$lmktval), sd(ceo_salary_df$lmktval), min(ceo_salary_df$lmktval), max(ceo_salary_df$lmktval))
ceoten. <- c(length(ceo_salary_df$ceoten), mean(ceo_salary_df$ceoten), sd(ceo_salary_df$ceoten), min(ceo_salary_df$ceoten), max(ceo_salary_df$ceoten))
ceotensq. <- c(length(ceo_salary_df$ceotensq), mean(ceo_salary_df$ceotensq), sd(ceo_salary_df$ceotensq), min(ceo_salary_df$ceotensq), max(ceo_salary_df$ceotensq))

salary_summary_df <- data.frame(lsales., lmktval., ceoten., ceotensq.)
row.names(salary_summary_df) <- c("COUNT", "MEAN", "STANDARD DEVIATION", "MINIMUM", "MAXIMUM")
salary_summary_df
```

#3c. RESTIMATION OF MODEL WITH WHITE STANDARD ERRORS AND THE T-STATS
```{r #3c. RESTIMATION OF MODEL WITH WHITE STANDARD ERRORS AND THE T-STATS}
coeftest(OLS_estimate_salary, vcov = sandwich)
coeftest(OLS_estimate_salary, vcov = vcovHC(OLS_estimate_salary, "HC0"))
coeftest(OLS_estimate_salary, vcov = vcovHC(OLS_estimate_salary)) 
summary(OLS_estimate_salary)
```

#3d. OBTAIN RESIDUALS AND STANDARDIZE
```{r #3d. OBTAIN RESIDUALS AND STANDARDIZE}
residuals <- residuals(OLS_estimate_salary)
residuals
standardised_residuals <- rstandard(OLS_estimate_salary)
standardised_residuals
```

#3e. RE-ESTIMATE MODEL BY ADDING DUMMY "COLLEGE" AND INTERACTION TERM "COLLEGE*LSALES". INTERPRETE COEFFICIENT OF INTERACTION TERM
```{r #3e. RE-ESTIMATE MODEL BY ADDING DUMMY "COLLEGE" AND INTERACTION TERM "COLLEGE*LSALES". INTERPRETE COEFFICIENT OF INTERACTION TERM}
OLS_reestimate_salary <- lm(lsalary ~ lsales + lmktval + ceoten + ceotensq + college + college*sales, data = ceo_salary_df)
summary(OLS_reestimate_salary)
```

#3f. RE-ESTIMATE MODEL WITH VARIABLE "LSALES_ADJUSTED" INSTEAD OF "LSALES". COMPARE COEFFICIENT ON "LSALES_ADJUSTED" WITH "LSALES"
```{r #3f. RE-ESTIMATE MODEL WITH VARIABLE "LSALES_ADJUSTED" INSTEAD OF "LSALES". COMPARE COEFFICIENT ON "LSALES_ADJUSTED" WITH "LSALES"}
ceo_salary_df$lsales_adjusted <- ceo_salary_df$lsales * 0.9
OLS_reestimate_salary2 <- lm(lsalary ~ lsales_adjusted + lmktval + ceoten + ceotensq + college + college*sales, data = ceo_salary_df)
summary(OLS_reestimate_salary2)

#The original coefficient of lsales  is less than the coefficient of lsales_adjusted.
```

## QUESTION 4 ##

# IMPORT DATA INTO R #
```{r # IMPORT DATA INTO R #}

library(haven)
library(sandwich)
library(foreign)
library(lmtest)
library(zoo)
library(dplyr)
library("plm")

local_returns_df <-read_dta("local_returns.dta")
```

#4a. CONFIRM IF BALANCED OR UNBALANCED PANEL. PROVIDE AGGREGATE STATISTICS
```{r #4a. CONFIRM IF BALANCED OR UNBALANCED PANEL. PROVIDE AGGREGATE STATISTICS}
local_returns_panel <- pdata.frame(local_returns_df, index = c("permno", "date"))
is.pbalanced(local_returns_panel)
aggregate(local_returns_panel, list(local_returns_df$year), summary)
```

#4b. REPORT AVERAGE RETURNS AND VOLATILITIES OF FIRMS IN THE 20 CITIES. CHECK FOR DIFFERENTIAL RETURNS AMONGST CITIES
```{r #4b. REPORT AVERAGE RETURNS AND VOLATILITIES OF FIRMS IN THE 20 CITIES. CHECK FOR DIFFERENTIAL RETURNS AMONGST CITIES}
aggregate(local_returns_panel, list(local_returns_df$city_returns), mean)
```

#4c. ESTIMATE PANEL REGRESSION
```{r}
local_returns_estimation <- lm(ret ~ city_returns + indret, data = local_returns_panel)
summary(local_returns_estimation)
```

#4d. ESTIMATE PANEL REGRESSION USING POOLED OLS WITHOUT DUMMIES
```{r #4d. ESTIMATE PANEL REGRESSION USING POOLED OLS WITHOUT DUMMIES}
local_returns_estimation_pooled <- plm(ret ~ city_returns + indret, data = local_returns_panel, model = "pooling")
summary(local_returns_estimation_pooled)
```

#4e. TEST FOR HETEROSKEDASTICITY IN THE RESIDUALS
```{r #4e. TEST FOR HETEROSKEDASTICITY IN THE RESIDUALS}
bptest(local_returns_estimation_pooled, data = local_returns_panel, studentize = F)

#Answer: The p-value is less than 5%. Hence, at 5% significance level we reject the null hypothesis and conclude that there is heteroskedasticity in the residuals.
```

#4f. ESTIMATE PANEL REGRESSION USING POOLED OLS WITH TIME FIXED EFFECTS (e.g. TIME DUMMIES)
```{r #4f. ESTIMATE PANEL REGRESSION USING POOLED OLS WITH TIME FIXED EFFECTS (e.g. TIME DUMMIES)}
```

#4g. ESTIMATE PANEL REGRESSION USING POOLED OLS WITH TIME AND FIRM FIXED EFFECTS. COMPARE RESULTS.
```{r #4g. ESTIMATE PANEL REGRESSION USING POOLED OLS WITH TIME AND FIRM FIXED EFFECTS. COMPARE RESULTS.}
library(Matrix)
library(lfe)
local_returns_estimation_pooled_results_bothtimeandfirm <- felm(ret ~ city_returns + indret | permno + year | 0 | 0, local_returns_panel)
summary(local_returns_estimation_pooled_results_bothtimeandfirm)
```

#4h. ESTIMATE PANEL REGRESSION USING POOLED OLS WITH TIME AND FIRM FIXED EFFECTS, AND WITH DOUBLE CLUSTERING ON BOTH FIRM AND TIME DIMENSIONS
```{r #4h. ESTIMATE PANEL REGRESSION USING POOLED OLS WITH TIME AND FIRM FIXED EFFECTS, AND WITH DOUBLE CLUSTERING ON BOTH FIRM AND TIME DIMENSIONS}
local_returns_estimation_pooled_results_bothtimeandfirm_clust <-felm(ret ~ city_returns + indret | permno + year | 0 | permno + year, data = local_returns_panel)
summary(local_returns_estimation_pooled_results_bothtimeandfirm_clust)
```

#4i. ESTIMATE PANEL REGRESSION USING FIXED EFFECT (FE) ESTIMATOR
```{r #4i. ESTIMATE PANEL REGRESSION USING FIXED EFFECT (FE) ESTIMATOR}
local_returns_estimation_firmfe <-felm(ret ~ city_returns + indret | permno | 0 | 0, data=local_returns_df)
summary(local_returns_estimation_firmfe)
```

#4j. ESTIMATE PANEL REGRESSION USING FIRST DIFFERENCE (FD) ESTIMATOR
```{r #4j. ESTIMATE PANEL REGRESSION USING FIRST DIFFERENCE (FD) ESTIMATOR}
local_returns_estimation_firstdiff <- plm(ret ~ city_returns + indret, data=local_returns_df, model = "fd")
  coef(summary(local_returns_estimation_firstdiff))
```
